An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-09-29 paper.tex
Rework Duration of Participation in §5.1

%..................................................................................................
\pb{Contribution duration.}
We now look into the longevity of contributors, in terms of the number of
years that they actively participate in any IETF mailing lists. We define
the \textit{contribution duration} of a participant as the length of time
that they have contributed to the mailing lists.  To do this, we look at
the contributors who first contribute to mailing lists between the years
2000 to 2013. We limit our analysis to 2013, since the longevity of
contributors that first participate more recently cannot be determined.
For each year, in the period 2000 to 2013, we look at those who first
contribute in each year and the number of years that they then go on to
remain active in the mailing lists. For example, a participant who first
sent an e-mail to an IETF list in 2010, and last sent an e-mail to an IETF
list in 2018, will have a contribution duration of 9 years.

We generate Gaussian Mixture Models for observing different clusters of the
maximum duration that contributors remain active for.  These models
identified three broad clusters that contributors can be categorised
within:

\begin{itemize}
  \item \emph{Young contributors} who leave within 1 year of joining.
  \item \emph{Mid-age contributors} who go onto remain associated for
    more than 1 year, but less than 5 years.
  \item \emph{Senior contributors} who go onto remain associated for 5 or
    more years.
\end{itemize}

We use these contribution categories defined above to characterise the
interactions between contributors belonging to different categories, as
well as the authors of each RFC. We look at outgoing and incoming
interactions with RFC authors in the period between the first draft and
publication of the RFC. If this period is less than two years, we look at
the activities of authors for two years before the RFC was published.
Interactions are defined from the viewpoint of the RFC authors:

\begin{itemize}
  \item \textit{Outgoing interaction}, where an RFC author responds
    to an email from other contributors (i.e., email sent).

  \item \textit{Incoming interaction}, where contributors respond to
    an email by the author (i.e., email received).
\end{itemize}

For each interaction type, we count the number of messages and number of
unique contributors involved in those interactions. For each RFC, we
analyse the interaction activities based on three measures of author
contribution duration:

\begin{itemize}
  \item \emph{Junior-most author}, i.e., the author with the lowest
    contribution duration at the time of publication.
  \item \emph{Senior-most author}, i.e., the author with the highest
    contribution duration at time of publication.
  \item \emph{Mean contribution duration} of all of the authors of the
    RFC at the time of publication.
\end{itemize}

@@

Figure~\ref{fig:age_dist_rfc_authors} shows the distribution of
contribution duration of each of these three measures. This shows that the
majority of junior-most authors have participated for less than 5 years in
the IETF, whereas the majority of senior-most authors have participated in
the IETF for more than 10 years. In fact, 35\% of authors exceed 15 years
of IETF participation, showing that RFCs tend to be authored by a mix of
seniority levels.

>>>>

%..................................................................................................
\pb{Duration of Participation:}
We now look into the duration of participation, in terms of the number of
years that people actively participate in any IETF mailing lists.

We define the \textit{contribution duration} of a participant as the length
of time that they have contributed to the mailing lists.  To do this, we
study the participants who first send email to mailing lists between the
years 2000 to 2013. We limit our analysis to 2013, since the longevity of
contributors that first participate more recently cannot be determined.
For each year, in the period 2000 to 2013, we look at those who first
sent email in each year and the number of years that they then go on to
remain active in the mailing lists. For example, a participant who first
sent an e-mail to an IETF list in 2010, and last sent an e-mail to an IETF
list in 2018, will have a contribution duration of 9 years.

First, to understand duration of participation, we first generate Gaussian
Mixture Models for observing different clusters of the maximum duration of
contribution. These models identified three broad clusters into which the
participants can be categorised: \emph{young contributors} for whom the
time between their first and last emails to the IETF lists in one year or
less; \emph{mid-age contributors} who go onto remain active participants in
the email lists for more than one year, but less than five years; and
\emph{senior contributors} who remain participants in the mailing list for
five or more years.

Next, to begin to understand how contribution duration relates to RFC
authorship, we look at the email interactions of the authors of each RFC.
Specifically, we look at the distribution of the contribution duration of
the authors of RFCs, considering the \emph{mean contribution duration} of
all of the authors of the RFC at the time of publication, the contribution
of the \emph{junior-most author} (the author with the lowest contribution
duration at the time of publication), and the contribution of the
\emph{senior-most author} (the author with the highest contribution
duration at time of publication.  Figure~\ref{fig:age_dist_rfc_authors}
shows the distribution of contribution duration of each of these three
measures. This shows that the majority of junior-most authors have
participated for less than 5 years in the IETF (i.e., they are young or
mid-age contributors), whereas the majority of senior-most authors are
senior contributors who have participated in the IETF for significantly
more than 5 years (in fact, 35\% of authors exceed 15 years of IETF
participation).  This shows that RFCs tend to be authored by a mix of
seniority levels.

% We look at outgoing and incoming interactions with RFC authors in the
% period between the first draft and publication of the RFC. If this period
% is less than two years, we look at the activities of authors for two
% years before the RFC was published.  Interactions are defined from the
% viewpoint of the RFC authors: \emph{outgoing interactions} where an RFC
% author responds to an email from other contributors (i.e., email sent);
% and \emph{incoming interactions} where contributors respond to an email
% by the author (i.e., email received).

@@
